TODO LIST

### 2019-01-19 LMCS/sendrev/saluja.tex
correcciones en la carpeta sendrev
carta de respuesta + correcciones varias

Inspired by the connection between $\shp$ and $\sfo$, a hierarchy of subclasses of $\sfo$ was introduced in~\cite{SalujaST95} 
by restricting the alternation of quantifiers in Boolean formulae.
Specifically, the \emph{$\sfo$-hierarchy} consists of the 
the classes $\E{i}$ and $\U{i}$ for every $i \geq 0$, where $\E{i}$ (resp., $\U{i}$) is defined as $\sfo$ but restricting the formulae used to be in $\loge{i}$ (resp., $\logu{i}$).
By definition, we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
\[
\E{0} \; \subsetneq \; \E{1} \; \subsetneq \; \U{1} \; \subsetneq \; \E{2} \; \subsetneq \; \U{2} \; = \; \sfo 
\]
In light of the framework introduced in this paper, natural extensions of these classes are obtained by considering 
$\eqso(\loge{i})$ and $\eqso(\logu{i})$ for every $i \geq 0$, which form the \emph{$\eqso(\fo)$-hierarchy}.
Clearly, we have that $\E{i} \subseteq \QE{i}$ and $\U{i} \subseteq \QU{i}$. Indeed, each formula $\varphi(\bar{X}, \bar{x})$ in $\E{i}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\QE{i}$, and likewise for $\U{i}$ and $\QU{i}$.
But what is the exact relationship between these two hierarchies?
To answer this question, we first introduce two normal forms for $\eqso(\LL)$ that helps us to characterize the expressive power of this quantitative logic.
A formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-prenex normal form} ($\LL$-PNF) 
if $\alpha$ is of the form
$\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})$,
where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, (as expected, $\sa{\bar{X}}\!$ is the respective nesting of $\sa{X}$'s) and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$. Notice that 
a formula $\varphi(\bar{X}, \bar{x})$ in $\sh{\LL}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\LL$-PNF. 
Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form} ($\LL$-SNF) if $\alpha$ is of the form $\sum_{i=1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-PNF. 
\begin{prop}\label{theo-pnf-snf}
Every formula in $\eqso(\LL)$ can be rewritten in $\LL$-SNF.
\end{prop}
\proof
\input{proof-theo-pnf-snf}
\qed
If a formula is in $\LL$-PNF then clearly the formula is in $\LL$-SNF.
Unfortunately, for some $\LL$ there exist formulae in $\eqso(\LL)$  that cannot be rewritten in $\LL$-PNF.
Therefore, to unveil the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, we need to understand the boundary between PNF and SNF. We do this in the following theorem. 
\begin{thm}\label{theo-pi1-pnf}
For $i = 0,1$, there exists a formula $\alpha_i$ in $\QE{i}$ that is not equivalent to any formula in $\Sigma_i$-PNF. 
On the other hand, if $\logu{1} \subseteq \LL$ and $\LL$ is closed under conjunction and disjunction, then every formula in $\eqso(\LL)$ can be rewritten in $\LL$-PNF. 
\end{thm}
\proof
\input{proof-theo-pi1-pnf}
\qed

>>>>

Inspired by the connection between $\shp$ and $\sfo$, a hierarchy of subclasses of $\sfo$ was introduced in~\cite{SalujaST95} 
by restricting the alternation of quantifiers in Boolean formulae.
Specifically, the \emph{$\sfo$-hierarchy} consists of the 
the classes $\E{i}$ and $\U{i}$ for every $i \geq 0$, where $\E{i}$ (resp., $\U{i}$) is defined as $\sfo$ but restricting the formulae used to be in $\loge{i}$ (resp., $\logu{i}$).
By definition, we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
\[
\E{0} \; \subsetneq \; \E{1} \; \subsetneq \; \U{1} \; \subsetneq \; \E{2} \; \subsetneq \; \U{2} \; = \; \sfo 
\]
In light of the framework introduced in this paper, natural extensions of these classes are obtained by considering 
$\eqso(\loge{i})$ and $\eqso(\logu{i})$ for every $i \geq 0$, which form the \emph{$\eqso(\fo)$-hierarchy}.
Clearly, we have that $\E{i} \subseteq \QE{i}$ and $\U{i} \subseteq \QU{i}$. Indeed, each formula $\varphi(\bar{X}, \bar{x})$ in $\E{i}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\QE{i}$, and likewise for $\U{i}$ and $\QU{i}$.
But what is the exact relationship between these two hierarchies?
To answer this question, we first introduce two normal forms for $\eqso(\LL)$ that helps us to characterize the expressive power of this quantitative logic.
A formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-prenex normal form} ($\LL$-PNF) 
if $\alpha$ is of the form
$\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})$,
where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, (as expected, $\sa{\bar{X}}\!$ and $\sa{\bar{x}}\!$ are the respective nestings of $\sa{X}$'s and $\sa{x}$'s) and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$. Notice that 
a formula $\varphi(\bar{X}, \bar{x})$ in $\sh{\LL}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\LL$-PNF. 
Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form} ($\LL$-SNF) if $\alpha$ is of the form $\sum_{i=1}^n \alpha_i$, where this is a shorthand notation for $\alpha_1+\cdots+\alpha_n$, and each $\alpha_i$ is in $\LL$-PNF. 
\begin{prop}\label{theo-pnf-snf}
Every formula in $\eqso(\LL)$ can be rewritten in $\LL$-SNF.
\end{prop}
\proof
\input{proof-theo-pnf-snf}
\qed
If a formula is in $\LL$-PNF then clearly the formula is in $\LL$-SNF.
Unfortunately, for some $\LL$ there exist formulae in $\eqso(\LL)$  that cannot be rewritten in $\LL$-PNF.
Therefore, to unveil the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, we need to understand the boundary between PNF and SNF. We do this in the following theorem. 
\begin{thm}\label{theo-pi1-pnf}
For $i = 0,1$, there exists a formula $\alpha_i$ in $\QE{i}$ that is not equivalent to any formula in $\Sigma_i$-PNF. 
On the other hand, if $\logu{1} \subseteq \LL$ and $\LL$ is closed under conjunction and disjunction, then every formula in $\eqso(\LL)$ can be rewritten in $\LL$-PNF. 
\end{thm}
\proof
\input{proof-theo-pi1-pnf}
\qed

### 2019-01-19 LMCS/sendrev/todo-list.tex
carta de respuesta + correcciones varias

% !TeX root = main.tex

\section*{\LARGE \textbf{TODO LIST}}

{\bf Reviewer 1}

{\bf Technical}

@@

\item p.2, first paragraph: ``However, other function classes have emerged from the need to understand

@@

\item p.7, first sentence: ``fragments or extensions [...] obtained by restricting'': extensions are not

@@

\item p.10, before Thm. 4.5: ``by considering second-order product'': We need both second-order

@@

\item p.10, last line: ``$B$-th bit'': mention in what way $B$ is interpreted as a number

@@

\item p.14: ``we have to concentrate on the class of counting problems with easy decision versions'':

@@

\item p.14: ``as expected, $\Sigma \bar{X}$ is the respective nesting of $\Sigma X.$’s'': why is this explained for secondorder

@@

\item p.14: $\sum_{i=1}^n\alpha_i$: explain that this is notation for $\alpha_1+\cdots+\alpha_n$
\item p.17: ``Since $\alpha$ is not null'' $\to$ ``Since $\sem{\alpha}$ is not the identically zero function'' (or similar)

@@

\item p.18, after proof of Claim 5.4: ``which implies that 2 $\in \eqso(\Sigma_0)$ and $2 \not\in \E{1}$'': $2 \in \eqso(\Sigma_0)$

@@

\item p.20: ``Since $\Sigma_1$ is closed under conjunction, then Lemma 5.7 holds for ...'' $\to$ ``Since $\Sigma_1$ is

@@

\item p.22: ``To show that $\alpha$ is in \totp'' $\to$ ``To show that the function defined by $\alpha$ ...'' (or similar)

@@

{\bf Grammatical/Language}

@@

{\bf Grammatical/Language}

@@

{\bf Reviewer 2}

@@

\item[$\checkmark$] page 4:
-some of the $\to$ some of them 
\item[$\checkmark$] page 10: 
-$2^{|A|^l}$-bits $\to$ $2^{|A|^l}$ bits
\item[$\checkmark$] page 21:
-it does not mentioned $\to$ it does not mention
\item page 22: 
-you make a long proof about some closure under subtraction by one.
You should put the result in a theorem envirpnment to make what you are
proving more evident.
\item[$\checkmark$] -replace .... for $\to$ replace ... by;  there are several instances of
that
\item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
\item -you explain that you rewrite your formula to obtain conditions a,b and c, 
you should give an idea how these conditions will be used
\item[$\checkmark$] page 23:
a SO, a FO  an SO, an FO
\item page 24: 
if, if $\phi$$\to$ $\phi$
\item page 25: 
-you explain that the proof relies on witnesses of logarithmic size.
It was not clear to me while reading the proof. Can you explain more ?
\item page 28:
-the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn. You should mention here that you do not need sum here
and that the framework of [39] is enough for this result.
\item[?] page 30:
-not, trivial $\to$ not trivial
\item[$\checkmark$] page 32: 
- paths of length at most n $\to$ in an acylic graph, all paths are of size less than n
\item page 33:
-there is a proof symbol \verb|\qed| but it does not end a well defined proof environment

@@

>>>>

\documentclass[a4paper]{article}

\usepackage{fullpage}
\usepackage{cite}
\usepackage{amsmath,amsfonts,amssymb,mathabx}
\usepackage{stmaryrd}
\usepackage[textwidth=2cm,textsize=small]{todonotes}
\usepackage{enumitem}
\input{macros}

\def\eg{{\em e.g.}}
\def\cf{{\em cf.}}

\begin{document}

\title{TODO LIST}
\date{}

\maketitle

\section*{Reviewer 1}

\subsection*{Technical}

@@

\item[$\checkmark$] p.2, first paragraph: ``However, other function classes have emerged from the need to understand

@@

\item[$\checkmark$] p.7, first sentence: ``fragments or extensions [...] obtained by restricting'': extensions are not

@@

\item[$\checkmark$] p.10, before Thm. 4.5: ``by considering second-order product'': We need both second-order

@@

\item[$\checkmark$] p.10, last line: ``$B$-th bit'': mention in what way $B$ is interpreted as a number

@@

\item[$\checkmark$] p.14: ``we have to concentrate on the class of counting problems with easy decision versions'':

@@

\item[$\checkmark$] p.14: ``as expected, $\Sigma \bar{X}$ is the respective nesting of $\Sigma X.$’s'': why is this explained for secondorder

@@

\item[$\checkmark$] p.14: $\sum_{i=1}^n\alpha_i$: explain that this is notation for $\alpha_1+\cdots+\alpha_n$
\item[$\checkmark$] p.17: ``Since $\alpha$ is not null'' $\to$ ``Since $\sem{\alpha}$ is not the identically zero function'' (or similar)

@@

\item[$\checkmark$] p.18, after proof of Claim 5.4: ``which implies that 2 $\in \eqso(\Sigma_0)$ and $2 \not\in \E{1}$'': $2 \in \eqso(\Sigma_0)$

@@

\item[$\checkmark$] p.20: ``Since $\Sigma_1$ is closed under conjunction, then Lemma 5.7 holds for ...'' $\to$ ``Since $\Sigma_1$ is

@@

\item[$\checkmark$] p.22: ``To show that $\alpha$ is in \totp'' $\to$ ``To show that the function defined by $\alpha$ ...'' (or similar)

@@

\subsection*{Grammatical/Language}

@@

\subsection*{Typos}

@@

\section*{Reviewer 2}

@@

\item[$\checkmark$] page 4:
-some of the $\to$ some of them 
\item[$\checkmark$] page 10: 
-$2^{|A|^l}$-bits $\to$ $2^{|A|^l}$ bits
\item[$\checkmark$] page 21:
-it does not mentioned $\to$ it does not mention
\item page 22: 
-you make a long proof about some closure under subtraction by one.
You should put the result in a theorem envirpnment to make what you are
proving more evident.
\item[$\checkmark$] -replace .... for $\to$ replace ... by;  there are several instances of
that
\item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
\item -you explain that you rewrite your formula to obtain conditions a,b and c, 
you should give an idea how these conditions will be used
\item[$\checkmark$] page 23:
a SO, a FO  an SO, an FO
\item page 24: 
if, if $\phi$$\to$ $\phi$
\item page 25: 
-you explain that the proof relies on witnesses of logarithmic size.
It was not clear to me while reading the proof. Can you explain more ?
\item page 28:
-the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn. You should mention here that you do not need sum here
and that the framework of [39] is enough for this result.
\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''

\item[$\checkmark$] page 32: 
- paths of length at most n $\to$ in an acylic graph, all paths are of size less than n
\item page 33:
-there is a proof symbol \verb|\qed| but it does not end a well defined proof environment

@@

\end{document}
